---
title: "Replication: A Contribution to the Empirics of Economic Growth by Mankiw, Romer, and Weil (1992)"
output: html_document
---

## R Libraries and dataset

```{r message=FALSE}
library(tibble)     # nice dataframes
library(stargazer)  # regression tables
library(ggplot2)    # nice graphs
library(dplyr)      # data manipulation
library(car)        # test linear hypotheses
library(skimr)      # descriptive statistics
library(haven)      # reading stata data

options(warn = -1)
options(scipen = 10000)
options(repr.plot.width = 6, repr.plot.height = 4)

mrw_data <- read_dta("https://github.com/freddyacuna/ECONOMETRICS-I-UC/blob/master/Undergraduate/2021%20Spring/HW2/mrw.dta")

```

## Intro 

```{r}
skim(mrw_data)

```
Rename variable names, so they are more meaningful.

```{r}
mrw_data <- mrw_data %>% 
              rename(non_oil = n, 
                     oecd = o,
                     intermediate = i,
                     gdp_60 = rgdpw60,
                     gdp_85 = rgdpw85,
                     gdp_growth_60_85 = gdpgrowth,
                     pop_growth_60_85 = popgrowth,
                     inv_gdp = i_y,
                     school = school)
```


Create the key variables of the Solow model.

```{r}
delta_gamma <- 0.05

mrw_data <- mrw_data %>% 
                  mutate(ln_gdp_85 = log(gdp_85),
                         ln_gdp_60 = log(gdp_60),
                         ln_gdp_growth = ln_gdp_85 - ln_gdp_60,
                         ln_inv_gdp = log(inv_gdp/100),
                         non_oil = factor(non_oil),
                         intermediate = factor(intermediate),
                         oecd = factor(oecd),
                         ln_ndg = log(pop_growth_60_85/100 + delta_gamma),
                         ln_school = log(school/100)) %>% 
                  select(country, region, ln_gdp_85, ln_gdp_60, ln_inv_gdp, 
                         non_oil, intermediate, oecd,
                         ln_ndg, ln_school, gdp_growth_60_85, ln_gdp_growth)
```




```{r}
skim(mrw_data)
```




## Data samples

```{r}
non_oil      <- mrw_data %>% 
                    filter(non_oil == 1)
intermediate <- mrw_data %>% 
                    filter(intermediate == 1)
oecd         <- mrw_data %>% 
                    filter(oecd == 1)
```
Additional samples for further exploration

```{r}
africa       <- mrw_data %>% 
                    filter(region == "Africa")
asia         <- mrw_data %>% 
                    filter(region == "Asia")
europe       <- mrw_data %>% 
                    filter(region == "Europe")
latinAmerica <- mrw_data %>% 
                    filter(region == "LatinAmerica")
northAmerica <- mrw_data %>% 
                    filter(region == "NorthAmerica")
oceania      <- mrw_data %>% 
                    filter(region == "Oceania")
```

## Solow model: Unrestricted Regression

```{r}
solow_oecd <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg, data = oecd)
solow_int  <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg, data = intermediate)
solow_noil <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg, data = non_oil)
```


```{r}
stargazer(solow_noil, solow_int, solow_oecd, digits=2, type = "text",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(I / GDP)", 
                               "log(n+delta+g)", 
                               "Constant"), 
          dep.var.labels = "Log(GDP) 1985",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 1 - Unrestricted Models",
          style = "qje")
```
## Solow model: Restricted  Regression

```{r}
solow_noil_r <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg), data = non_oil)
solow_int_r  <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg), data = intermediate)
solow_oecd_r <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg), data = oecd)
```



```{r}
stargazer(solow_noil_r, solow_int_r, solow_oecd_r, digits=2, type = "text",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(I / GDP)- log(n+delta+g)", 
                               "Constant"), 
          dep.var.labels = "Log(GDP) 1985",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 1 - Restricted Models",
          style = "qje")
```
## Solow model: Test of Restriction

$H0:$ The coefficients on $\ln (s)$  and $\lnâ¡(n+\gamma+\delta)$ are equal in magnitude and opposite in sign.

```{r}
linearHypothesis(solow_noil, "ln_inv_gdp = - ln_ndg")
```


```{r}
linearHypothesis(solow_int,  "ln_inv_gdp = - ln_ndg")
```



```{r}
linearHypothesis(solow_oecd, "ln_inv_gdp = - ln_ndg")
```

## Solow model: Implied alpha


```{r}
C <- coef(solow_noil_r)[2]
alpha_solow_noil_r <- C/(1+C)
alpha_solow_noil_r <- round(alpha_solow_noil_r, 2)
print(paste("Implied alpha (Non oil):", alpha_solow_noil_r))
```


```{r}
C <- coef(solow_int_r)[2]
alpha_solow_int_r <- C/(1+C)
alpha_solow_int_r <- round(alpha_solow_int_r, 2)
print(paste("Implied alpha (Intermediate):", alpha_solow_int_r))
```



```{r}
C <- coef(solow_oecd_r)[2]
alpha_solow_oecd_r <- C/(1+C)
alpha_solow_oecd_r <- round(alpha_solow_oecd_r, 2)
print(paste("Implied alpha (OECD):", alpha_solow_oecd_r))
```


## Augmented Solow model: Unrestricted regression

```{r}
augsolow_noil <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg + ln_school, data = non_oil)
augsolow_int  <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg + ln_school, data = intermediate)
augsolow_oecd <- lm(ln_gdp_85 ~ ln_inv_gdp + ln_ndg + ln_school, data = oecd)
```


```{r}
stargazer(augsolow_noil, augsolow_int, augsolow_oecd, digits=2, type = "text",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(I / GDP)", 
                               "log(n+delta+g)",
                               "log(school)",
                               "Constant"), 
          dep.var.labels = "Log(GDP) 1985",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 2 - Unrestricted Models",
          style = "qje")
```





## Augmented Solow model: Restricted regression

```{r}
augsolow_noil_r <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = non_oil)
augsolow_int_r  <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = intermediate)
augsolow_oecd_r <- lm(ln_gdp_85 ~ I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = oecd)
```

```{r}
stargazer(augsolow_noil_r, augsolow_int_r, augsolow_oecd_r, digits=2, type = "text",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(I / GDP)- log(n+delta+g)", 
                               "log(school)- log(n+delta+g)",
                               "Constant"), 
          dep.var.labels = "Log(GDP) 1985",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 2 - Restricted Models",
          style = "qje")
```

## Augmented Solow model: Test of restriction

```{r}
linearHypothesis(augsolow_oecd, "ln_inv_gdp + ln_ndg + ln_school = 0")
```



```{r}
linearHypothesis(augsolow_noil, "ln_inv_gdp + ln_ndg + ln_school = 0")
```

```{r}
linearHypothesis(augsolow_int, "ln_inv_gdp + ln_ndg + ln_school = 0")
```






## Augmented Solow model: Implied alpha and beta



```{r}
c2 <- round(coef(augsolow_noil_r)[2], 2)
c2
c3 <- round(coef(augsolow_noil_r)[3], 2)
c3
```

```{r}
x = c2 / (c2+c3+1)
x = round(x, 2)
print(paste("Implied alpha (Non oil):", x))
y = c3 / (c2+c3+1)
y = round(y, 2)
print(paste("Implied beta (Non oil):", y))
```


```{r}
c2 <- round(coef(augsolow_int_r)[2], 2)
c2
c3 <- round(coef(augsolow_int_r)[3], 2)
c3
```



```{r}
x = c2 / (c2+c3+1)
x = round(x, 2)
print(paste("Implied alpha (Intermediate):", x))
y = c3 / (c2+c3+1)
y = round(y, 2)
print(paste("Implied beta (Intermediate):", y))
```



```{r}
c2 <- round(coef(augsolow_oecd_r)[2], 2)
c2
c3 <- round(coef(augsolow_oecd_r)[3], 2)
c3
```


```{r}
x = c2 / (c2+c3+1)
x = round(x, 2)
print(paste("Implied alpha (OECD):", x))
y = c3 / (c2+c3+1)
y = round(y, 2)
print(paste("Implied beta (OECD):", y))
```

## Unconditional convergence
```{r}
ucc_noil <- lm(ln_gdp_growth ~ ln_gdp_60, data = non_oil)
ucc_int  <- lm(ln_gdp_growth ~ ln_gdp_60, data = intermediate)
ucc_oecd <- lm(ln_gdp_growth ~ ln_gdp_60, data = oecd)
```



```{r}
stargazer(ucc_noil, ucc_int, ucc_oecd, digits=2, type = "text",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(GDP '60)",
                              "Constant"),
          dep.var.labels = "Log(GDP '85) - Log(GDP '60)",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 3 - Unconditional Convergence",
          style = "qje")
```

## Implied speed of convergence

```{r}
# Implied speed of convergence and halflife
speed = -log(1+coef(ucc_noil)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 5)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (Non-oil):", speed))
print(paste("Implied halflife of unconditional (Non-oil):", halfLife))
```



```{r}
# Implied speed of convergence and halflife
speed = - log(1+coef(ucc_int)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 5)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (Intermediate):", speed))
print(paste("Implied halflife of unconditional (Intermediate):", halfLife))
```



```{r}
# Implied speed of convergence and halflife
speed = - log(1+coef(ucc_oecd)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 4)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (OECD):", speed))
print(paste("Implied halflife of unconditional (OECD):", halfLife))
```

## Conditional convergence

```{r}
cc_noil <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg, data = non_oil)
cc_int  <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg, data = intermediate)
cc_oecd <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg, data = oecd)
```


```{r}
stargazer(cc_noil, cc_int, cc_oecd, digits=2, type = "text",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(GDP '60)",
                               "log(I / GDP)", 
                               "log(n+delta+g)",
                               "Constant"),
          dep.var.labels = "Log(GDP '85) - Log(GDP '60)",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 4 - Conditional Convergence",
          style = "qje")
```

## Implied speed of convergence

```{r}
# Implied speed of convergence and halflife
speed = -log(1+coef(cc_noil)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 5)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (Non-oil):", speed))
print(paste("Implied halflife of unconditional (Non-oil):", halfLife))
```

```{r}
# Implied speed of convergence and halflife
speed = - log(1+coef(cc_int)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 4)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (Intermediate):", speed))
print(paste("Implied halflife of unconditional (Intermediate):", halfLife))
```

```{r}
# Implied speed of convergence and halflife
speed = -log(1+coef(cc_oecd)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 4)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (OECD):", speed))
print(paste("Implied halflife of unconditional (OECD):", halfLife))
```

## Augmented conditional convergence
```{r}
augcc_noil <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg + ln_school, data = non_oil)
augcc_int  <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg + ln_school, data = intermediate)
augcc_oecd <- lm(ln_gdp_growth ~ ln_gdp_60 + ln_inv_gdp + ln_ndg + ln_school, data = oecd)
```

```{r}
stargazer(augcc_noil, augcc_int, augcc_oecd, digits=2, type = "text",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(GDP '60)",
                               "log(I / GDP)", 
                               "log(n+delta+g)",
                               "log(school)",
                               "Constant"),
          dep.var.labels = "Log(GDP '85) - Log(GDP '60)",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 5 - Augmented Conditional Convergence",
          style = "qje")
```

## Implied speed of convergence
```{r}
# Implied speed of convergence and halflife
speed = -log(1+coef(augcc_noil)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 4)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (Non-oil):", speed))
print(paste("Implied halflife of unconditional (Non-oil):", halfLife))
```

```{r}
# Implied speed of convergence and halflife
speed = - log(1+coef(augcc_int)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 4)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (Intermediate):", speed))
print(paste("Implied halflife of unconditional (Intermediate):", halfLife))
```

```{r}
# Implied speed of convergence and halflife
speed = -log(1+coef(augcc_oecd)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 4)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (OECD):", speed))
print(paste("Implied halflife of unconditional (OECD):", halfLife))
```

## Restricted augmented conditional convergence

```{r}
augcc_noil_r <- lm(ln_gdp_growth ~ ln_gdp_60 + I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = non_oil)
augcc_int_r  <- lm(ln_gdp_growth ~ ln_gdp_60 + I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = intermediate)
augcc_oecd_r <- lm(ln_gdp_growth ~ ln_gdp_60 + I(ln_inv_gdp - ln_ndg) + I(ln_school - ln_ndg), data = oecd)
```

```{r}
stargazer(augcc_noil_r, augcc_int_r, augcc_oecd_r, digits=3, type = "text",
          column.labels = c("Non-Oil", 
                            "Intermediate", 
                            "OECD"),
          covariate.labels = c("log(GDP '60)",
                               "log(I / GDP) - log(n+delta+g)", 
                               "log(school) - log(n+delta+g)",
                               "Constant"),
          dep.var.labels = "Log(GDP '85) - Log(GDP '60)",
          omit.stat = c("f", 
                        "rsq", 
                        "ser"),
          title = "Table 6 - Restricted Augmented Conditional Convergence",
          style = "qje")
```

## Test of restriction

```{r}
linearHypothesis(augcc_noil, "ln_inv_gdp + ln_ndg + ln_school = 0")
```

```{r}
linearHypothesis(augcc_int, "ln_inv_gdp + ln_ndg + ln_school = 0")
```

```{r}
linearHypothesis(augcc_oecd, "ln_inv_gdp + ln_ndg + ln_school = 0")
```

## Implied speed of convergence
```{r}
# Implied speed of convergence and halflife
speed = -log(1+coef(augcc_noil_r)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 4)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (Non-oil):", speed))
print(paste("Implied halflife of unconditional (Non-oil):", halfLife))
```

```{r}
# Implied speed of convergence and halflife
speed = -log(1+coef(augcc_int_r)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 4)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (Intermediate):", speed))
print(paste("Implied halflife of unconditional (Intermediate):", halfLife))
```


```{r}
# Implied speed of convergence and halflife
speed = -log(1+coef(augcc_oecd_r)[2])/(1985-1960)
halfLife = log(2)/speed

speed = round(speed, 4)
halfLife = round(halfLife)

print(paste("Implied speed of unconditional (OECD):", speed))
print(paste("Implied halflife of unconditional (OECD):", halfLife))
```
## Figures: Unconditional vs conditional convergence
```{r}
ggplot(intermediate, aes(x = ln_gdp_60, y = ln_gdp_growth)) +
    geom_point(shape = 1) +
    geom_smooth(method=lm, se=FALSE, color = "red") +
    theme_bw() +
    ggtitle("A: Unconditional") +
    ylab("Log Growth rate: 1960 - 85") +
    xlab("Log output per working age adult: 1960")
```
```{r}
y2 <- lm(ln_gdp_growth ~  ln_inv_gdp + ln_ndg, data = intermediate)$residuals
x2 <- lm(ln_gdp_60 ~  ln_inv_gdp + ln_ndg, data = intermediate)$residuals

panel_b <- tibble(y2, x2)
```
```{r}
ggplot(panel_b, aes(x = x2, y = y2)) +
    geom_point(shape = 1) +
    geom_smooth(method=lm, se=FALSE, color = "red") +
    theme_bw() +
    ggtitle("B: Conditional on Saving and Population Growth") +
    ylab("Log Growth rate: 1960 - 85") +
    xlab("Log output per working age adult: 1960")
```

```{r}

y3 <- lm(ln_gdp_growth ~  ln_inv_gdp + ln_ndg + ln_school, data = intermediate)$residuals
x3 <- lm(ln_gdp_60 ~  ln_inv_gdp + ln_ndg + ln_school, data = intermediate)$residuals

panel_c <- tibble(y3, x3)
```
```{r}
ggplot(panel_c, aes(x = x3, y = y3)) +
    geom_point(shape = 1) +
    geom_smooth(method=lm, se=FALSE, color = "red") +
    theme_bw() +
    ggtitle("C: Conditional on Saving and Population Growth and Human Capital") +
    ylab("Log Growth rate: 1960 - 85") +
    xlab("Log output per working age adult: 1960")
```
